---
title: "BIOL 471 Lecture 4"
author: "Kirt Onthank"
date: "04/02/2019"
output: pdf_document
---
#What does the 'print' function do?
```{r}
print(23)
```

# Basic loop function
```{r}
for (i in 10:20){
  print(i)
}
```


# Does not need to be sequential sequence. Can be arbitrary
```{r}
for (i in c(-3,10,1001,3,45,0.6)){
  print(i)
}
```


# Doesn't even need to be numbers
## You can also fill vectors before you call the loop
```{r}
thing=c("people","brown dog","fritos","tank","purple","thirty-six")

for (i in thing){
  print(i)
}
```

# What the 'rnorm' function does
```{r}
rnorm(10,mean=20,sd=7)
```

# Using i in a series of functions
```{r}
for (i in 20:40){
  x=rnorm(100,mean=i,sd=7)
  y=mean(x)
  z=sd(x)
  print(paste(i,y,z))
}
```

# Initialize a vector to store your output
```{r}
rmean=numeric()
rsd=numeric()

for (i in 1:20){
  x=rnorm(100,mean=i,sd=7)
  rmean[i]=mean(x)
  rsd[i]=sd(x)
}
rsd
```

# Another way to specify position in the 
```{r}
rmean=numeric()
rsd=numeric()
iter=1

for (i in c(-100,100,67,10000,0.5,10)){
  x=rnorm(100,mean=i,sd=7)
  rmean[iter]=mean(x)
  rsd[iter]=sd(x)
  iter=iter+1
}
rmean
rsd
```


# Be careful of how much time something takes
```{r}
time=numeric()
for(i in 1:8){
   start=Sys.time()
   x=rnorm(10^i,mean=100,sd=15)
   mean(x)
   time[i]=(Sys.time()-start)
}
plot(1:8,time)
```

# What does the 'sample' function do?
```{r}
sample(thing,1)
```

# Simulating 10 dice rolls
```{r}
dice=numeric()
for(i in 1:10){
dice[i]=sample(1:6,1)+sample(1:6,1)
print(dice)
}

```

# Just how fast is this function?
```{r}
start=Sys.time()
for(i in 1:50000){
dice[i]=sample(1:6,1)+sample(1:6,1)
}
Sys.time()-start
```

