---
title: "Lecture 22"
output: pdf_document
---

# Loading Packages
```{r Load Packages, results="hide"}
library(ape)
library(phangorn)
library(ips)
```

# Reading in Data
```{r Reading in FASTA}
bears=read.dna('bear_cytb.fasta',format="fasta")
```

# Neighbor Joining Tree

```{r Distance matrix}
bears.dist=dist.dna(bears)
bears.dist
```

```{r  Neightbor-joining tree, results='hide',fig.show="hide"}
bears.nj=NJ(bears.dist)
plot(bears.nj)
```



# Maximum Parsimony Tree

```{r Convert to phyDat}
bears.pd=as.phyDat(bears)
```

## Parsimony Tree Estimation

```{r Parsimony Tree,results="hide"}
treeMP=pratchet(bears.pd)
```

```{r}
plot(treeMP)
```

## Tree Branch support through bootstrapping

```{r Bootstrapping Tree, results='hide'}
BStrees=bootstrap.phyDat(bears.pd, pratchet, bs = 100)
```

```{r}
BStrees
```


```{r Plotting boot,fig.show=F,results=F,eval=F}
treeMP=plotBS(treeMP, BStrees, "phylogram")
```


```{r Re-rooting Tree}
treeMP.root=root(treeMP,"Ringed_Seal")
```

Now, lets re-plot your tree.
```{r Reploting tree with new root, fig.show=F,eval=F}
plotBS(treeMP.root, BStrees, "phylogram")
```

# Maximum Likelihood

## Selecting a Model of Nucleotide Substitution
```{r modelTest, results="hide"}
bears.mT=modelTest(bears.pd,bears.nj)
```

Lets take a look at the resulting table
```{r Model Table, results='hide'}
bears.mT
```

 Lets find the model with the lowest AICc value and use that. 

```{r Sorting Models, results='hide'}
bears.mT=bears.mT[order(bears.mT$AICc),]
bears.mT$Model[1]
```



## Substitution rate matrix
```{r Rate Matrix, results='hide'}
env=attr(bears.mT, "env")
model=get("HKY+G+I", env) # a call
eval(model, env=env)
```



```{r, results='hide',tidy=T}
fit=pml(bears.nj,bears.pd,model="HKY",optInv=T,optGamma=T)
```

```{r, results='hide', tidy=T}
fit=optim.pml(fit,model="HKY",optInv=T,optGamma=T,optNni=T)
```

```{r,eval=F,tidy=T}
plot(fit)
```


## Bootstraping

```{r,results='hide'}
bs=bootstrap.pml(fit, bs=100, optNni=TRUE)
```

```{r, eval=F}
treeML=plotBS(fit$tree,bs,type="phylogram")
```

## Rerooting tree
```{r,eval=F}
treeML.root=root(treeML,"Ringed_Seal")
plotBS(treeML.root,type="phylogram")
```


```{r,eval=F,tidy=T}
comparePhylo(treeML.root,treeMP.root,plot=T,force.rooted = T)
```


# Bayesian Tree with MrBayes

```{r,results='hide'}
mrbayes(bears, file="bears.nex",lset=mrbayes.lset(nst=2, rates="invgamma"),
        mcmc=mrbayes.mcmc(ngen=30000,printfreq = 1000,samplefreq = 100,
        nruns=2,nchains=2))
```


begin mrbayes;
	lset nst=6 rates=invgamma;
	mcmc ngen=300000 printfreq=10000 samplefreq=1000 nruns=2 nchains=2;
  sump filename=bears.nex burnin=0.25;
	sumt filename=bears.nex burnin=0.25 contype=allcompat conformat=simple;
end;



```{bash,eval=F}
mb bears.nex
```



```{r,eval=F}
mrbayes.tree=read.nexus("bears.nex.con.tre",tree.names=NULL)
```
.

```{r,eval=F}
plot(mrbayes.tree, show.node.label=T,cex=0.7)
```



```{r,eval=F,tidy=T}
comparePhylo(treeML.root,mrbayes.tree,plot=T,force.rooted = T)
```




