---
title: 'Lecture 10: Point Pattern Analysis'
author: "Kirt L Onthank"
date: "4/19/2020"
output: pdf_document
---

# Loading Libraries
```{r}
library(spatstat)
library(vcd)
set.seed(56)
```

# Quadrat Chi-squared Test for Spatial Randomness
## Generating a clustered spatial dataset and plotting
```{r}
lamb <-100/(90*90)
child <- 3
radius <- 2
pp.MAT<-rMatClust(kappa=(lamb/child),r=radius, mu=child,
                  win=owin(c(0,90),c(0,90)))
plot(pp.MAT,pch=21,bg="grey")
```


## Making 25 quadrats and counting points in the quadrats
```{r}
counts=quadratcount(pp.MAT, nx=5, ny=5)
plot(pp.MAT,pch=21,bg="grey")
plot(counts,add=T,cex=2)
```

## Chi-square test on quadrat counts
```{r}
chisq.test(counts)
```

## Showing the effects of quadrat size. Making 4 quadrats
```{r}
chisq.test(quadratcount(pp.MAT, nx=2, ny=2))
```

## Showing the effects of quadrat size. Making 64 quadrats
```{r}
chisq.test(quadratcount(pp.MAT, nx=15, ny=15))
```

## Generating a regular dataset
```{r}
pp.SSI<-rSSI(n=100, r=5, win = owin(c(0,90),c(0,90)),giveup=5000)
plot(pp.SSI,pch=21,bg="grey")
```

## Generating 25 quadrats and counts
```{r}
counts.reg=quadratcount(pp.SSI,nx=5,ny=5)
plot(pp.SSI,pch=21,bg="grey")
plot(counts.reg,add=T,cex=2)
```

## Chi-square test on regular counts
```{r}
chisq.test(counts.reg)
```

# Poisson Quadrat Test
```{r}
SSI.tab=table(quadratcount(pp.SSI,nx=5,ny=5))
SSI.tab
```

```{r}
SSI.gf=goodfit(SSI.tab,type= "poisson",method= "MinChisq", par=list(lambda=(mean(quadratcount(pp.SSI, nx=5, ny=5)))))
SSI.ma=as.matrix(SSI.gf)

summary(SSI.gf)
```

```{r}
plot(SSI.gf$count,SSI.gf$observed,col="red",lwd=2,xlab="Points/Quadrat",cex.lab=2,cex.axis=2,type="l",ylab="")
points(SSI.gf$count,SSI.gf$observed,pch=21,bg="red")
text(1,8,"Observed",col="red")
lines(SSI.gf$count,SSI.gf$fitted,col="green",lwd=2)
points(SSI.gf$count,SSI.gf$fitted,pch=21,bg="green")
text(1,5,"Expected",col="green")
```

# d1 neighborhood analysis
## d1 in Regular Data
```{r}
nearest=nndist(pp.SSI)
nearest.tab=table(nearest)

SSI.gf=goodfit(nearest.tab,type= "poisson",method= "MinChisq", par=list(lambda=(mean(quadratcount(pp.SSI, nx=5, ny=5)))))
summary (SSI.gf)
```

## d1 on clustered data
```{r}
nearest.mat=round(nndist(pp.MAT),1)
mat.nearest=table(nearest)

MAT.gf=goodfit(mat.nearest,type= "poisson",method= "MinChisq", par=list(lambda=(mean(quadratcount(pp.SSI, nx=5, ny=5)))))
summary (MAT.gf)
```

## Ripley's K and L test
### Ripley's K on clustered data
```{r}
plot(Kest(pp.MAT))
```

### Ripley's K on regular data
```{r}
plot(Kest(pp.SSI))
```

### L(t) function
```{r}
Lest2=function (...) { # the ... pass through definitions
  K=Kest(...) # straight to the Kest function
  nama <- colnames(K)
  K=K[, !(nama %in% c("rip", "ls"))]
  R=K$r
  L=eval.fv(R-sqrt(K/pi))
  L=rebadge.fv(L, substitute(L(r), NULL), "L")
  return(L) }
```

```{r}
SSI.lest=Lest2(pp.SSI)
plot(iso~r,data=Lest2(pp.SSI),type="l",col="red",ylab="L(t)",xlab="t",
     lwd=2,cex.axis=2,cex.lab=2,ylim=c(-5,5))
abline(h=0,lwd=2)
```


### Randomization Envelope Function
```{r}
lhat.env=function(n, s, r, win=owin(xrange=c(0,1), yrange=c(0,1))){
  hold=matrix(0, s, length(r))
  for(i in 1:s){
    hold[i,]=Lest2(runifpoint(n, win=win), r=r)$iso
  }
  mn=apply(hold, 2, mean)
  Up=apply(hold, 2, max)
  Down=apply(hold, 2, min)
  return(data.frame(mn, Up, Down))
}
```

### Running envelope function to determine significance
```{r}
SSI.env=lhat.env(100, 100, SSI.lest$r, win=owin(xrange=c(0,90), yrange=c(0,90)))
plot(iso~r,data=Lest2(pp.SSI),type="l",col="red",ylab="L(t)",xlab="t",
     lwd=2,cex.axis=2,cex.lab=2,ylim=c(-5,5))
lines(SSI.lest$r, SSI.env$Up, lty=2,lwd=2)
lines(SSI.lest$r, SSI.env$Down, lty=2,lwd=2)
lines(SSI.lest$r, SSI.env$mn, lty=2,lwd=2)
```

